---
title: "Codingchallenge6"
author: "Samit Kafle"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
```

**1.	2 pts. Regarding reproducibility, what is the main point of writing your own functions and iterations?**
Functions encapsulate specific tasks, making code modular and easier to understand, debug, and reuse.
Iterations (loops) allow us to apply the same operation across multiple elements, ensuring consistent processing.Functions and loops reduce redundancy, making code more efficient and less error-prone.


**2.	2 pts. In your own words, describe how to write a function and a for loop in R and how they work. Give me specifics like syntax, where to write code, and how the results are returned.** 
let's take example for calculating square of number
```{r}
#Define a function to calculate the square of a number*
#square <- function(x) {   This defines a function named square that takes one argument x
#  result <- x * x        This line calculates the square of x and stores it in result.
#  return(result)          This returns the value of result
#}

# Call the function
# square(4)  # Returns 16
```

```{r}
### For loop to calculate the square of each number
### Initialize an empty vector to store results
##squares <- c()

#for (i in numbers) {  This defines a loop that iterates over each element in the numbers                                     vector.
#squares <- c(squares, i * i)  This line calculates the square of i and appends it to the                                        squares vector.
#}

### Print the results
##print(squares)  # Returns c(1, 4, 9, 16, 25)
```



**3.	2 pts. Read in the Cities.csv file from Canvas using a relative file path.**
```{r}
datum=read.csv("Cities.csv")

```

**4.	6 pts. Write a function to calculate the distance between two pairs of coordinates based on the Haversine formula (see below). The input into the function should be lat1, lon1, lat2, and lon2. The function should return the object distance_km. All the code below needs to go into the function.**
```{r}
# Define the function to calculate distance using the Haversine formula
haversine_distance <- function(lat1, lon1, lat2, lon2) {
  # Convert to radians
  rad.lat1 <- lat1 * pi / 180
  rad.lon1 <- lon1 * pi / 180
  rad.lat2 <- lat2 * pi / 180
  rad.lon2 <- lon2 * pi / 180
  
  # Haversine formula
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a))
  
  # Earth's radius in kilometers
  earth_radius <- 6378137
  
  # Calculate the distance
  distance_km <- (earth_radius * c) / 1000
  
  # Return the distance
  return(distance_km)
}

```

**5.	5 pts. Using your function, compute the distance between Auburn, AL and New York City **
a.	Subset/filter the Cities.csv data to include only the latitude and longitude values you need and input as input to your function. 


```{r}
# Filter the data to include only Auburn, AL and New York City
filtered_data <- datum %>% 
  filter(city %in% c("Auburn", "New York"))

# Extract latitude and longitude values
lat1=datum[datum$city=="Auburn", "lat"]
lon1=datum[datum$city=="Auburn", "long"]
lat2=datum[datum$city=="New York", "lat"]
lon2=datum[datum$city=="New York", "long"]

```



b.	The output of your function should be 1367.854 km
```{r}
distance <- haversine_distance(lat1, lon1, lat2, lon2)
print(distance)
```
**6.	6 pts. Now, use your function within a for loop to calculate the distance between all other cities in the data. The output of the first 9 iterations is shown below. **
```{r}
# empty list to store distances
distances <- list()

# Iterate over each pair of cities in the data
for (i in 1:nrow(datum)) {
lat1=datum[datum$city=="Auburn", "lat"]
lon1=datum[datum$city=="Auburn", "long"]
lat2=datum[i, "lat"]
lon2=datum[i, "long"]
 
    
    # Calculate the distance using the function
    distance <- haversine_distance(lat1, lon1, lat2, lon2)
    
    # Append the result to the distances list
    distances <- append(distances, distance)
  }
print(distances)


```


**Bonus point if you can have the output of each iteration append a new row to a dataframe, generating a new column of data. In other words, the loop should create a dataframe with three columns called city1, city2, and distance_km, as shown below. The first six rows of the dataframe are shown below.**

```{r}
# Append the result to the dataframe
dist=data.frame(t(distances))# transforming the matrix to make it column
distance=t(dist)#
distances_list <- rbind(data.frame(City1 = "Auburn", City2 = datum$city, Distance_km = distance))
print(distances_list)
```


**7.	2 pts. Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 6. Provide me a link to your github written as a clickable link in your .pdf or .docx**

[link to the github](https://github.com/SamitKafle/codingchallenge6.git)


